---
title: "reading_mat"
author: "Xiaoni Xu"
date: "2024-11-18"
output: html_document
---

Loading needed packages
```{r, message = FALSE}
library(R.matlab)
library(ggplot2)
```




There are 5 shape metrics used: elastic scaled metric, elastic unscaled metric with elastic representation of shapes,  Full Procrustes metric, Kendall scaled metric, Kendall unscaled metric with Kendall's representation of shapes. 

Data in nested sturcture: factor-cell, animal, type of mitochondria (SS/IMF), exercise(Running/Sedentary)


### Reading `Allshape.mat`

`Allshape.mat` contains a large 3D array containing mitochondrial shape data for multiple animals and cells.

AllShape.mat contains all the reampled boundary, to make the code run faster we resample 50 points on the boundary.
%%%Data: 12 animal in total.X1,X2,...,X12 
%%%Each animal contain several cells: 4, 5 or 6.
%%%n1,n2,...,n12 = number of mitochondria shapes in each cell of each animal

```{r}
# Set the file path
Allshape <- "data/Allshape.mat"

# Read the .mat file
Allshape <- readMat(Allshape)
```

Try to visualize the mitochondrial shapes.
```{r}
# Extract mitochondria data for animal X1
mitochondria <- Allshape$X1

# Get the number of shapes for X1
num_shapes <- dim(mitochondria)[3]

# Set up a single plot layout for separate plots
par(mfrow = c(ceiling(sqrt(num_shapes)), ceiling(sqrt(num_shapes)))) # Square grid layout


# Set the number of plots per page (e.g., 16 plots per page)
plots_per_page <- 16
pages <- ceiling(num_shapes / plots_per_page)

for (page in 1:pages) {
  # Define the range of shapes to plot on this page
  start <- (page - 1) * plots_per_page + 1
  end <- min(page * plots_per_page, num_shapes)
  
  # Adjust layout for the current batch
  par(mfrow = c(4, 4), mar = c(2, 2, 2, 2)) # 4x4 grid with smaller margins
  
  # Plot each shape in the current batch
  for (i in start:end) {
    x_coords <- mitochondria[1, , i]
    y_coords <- mitochondria[2, , i]
    
    plot(x_coords, y_coords, type = "l", col = "blue",
         xlab = "", ylab = "", main = paste("Shape", i), asp = 1)
  }
  
  # Pause to view each page
  readline(prompt = "Press [Enter] to view the next page...")
}
```


Reading data.mat (same as Allshape.mat?)
```{r}
# Set the file path
data <- "data/data.mat"

# Read the .mat file
data <- readMat(data)

# Explore the structure of the loaded data
str(data)
```


### Reading classA.mat

Not sure what this is for - not included in the readme.txt file.
```{r}

# Set the file path
classA <- "data/classA.mat"

# Read the .mat file
classA <- readMat(classA)
```

### Reading A.mat

Not sure what this is for - not included in the readme.txt file.
```{r}

# Set the file path
A <- "data/A.mat"

# Read the .mat file
A <- readMat(A)
```

### Reading P_Elastic_scaled_cell.mat

```{r}

# Set the file path
P_Elastic_scaled_cell <- "data/P_Elastic_scaled_cell.mat"

# Read the .mat file
P_Elastic_scaled_cell <- readMat(P_Elastic_scaled_cell)
```

### Reading ScaledElasticDisMat.mat

ElasticDisMat.m

used to get the distance matrix between two groups/samples of shapes. 

To improve effciency, I use parallel computing with "parfor", it's just "for" looping.

```{r}
# Set the file path
ScaledElasticDisMat <- "data/ScaledElasticDisMat.mat"

# Read the .mat file
ScaledElasticDisMat <- readMat(ScaledElasticDisMat)
```



ReSampleCurve.m
Resample the original data to get N points on the boundary

plot_color_map.m
to display the distance matrix between two samples/groups

Ex: folder--ElasticShapeScaled


ElasticDistance.m
used to calculate the elastic scaled metric between two shapes/boundaries. 


EnergyStat2sample.m
used the calculate the energy statistic of two groups/samples

EnergyStatksample.m
used the calculate the energy statistic of k groups/samples

PermutationTest.m
use permutation test to get the p-value of energy test with energy statistic

TestStatistic.m
to test the significance of one factor when it contains another significant factor. 
For example, to test the the significance of animal when it contains another significant factor--cell.
I designed two test statistic: delta and epsilon but we think epsilon is better. 

TwoAnimalPermutation.m
use permutation test to get the p-value of of TestStatistic.m. Here, it's only for two-sample test 
because of the mitochondria data. If you have k-sample, you can design k-sample test statistic like 
energy statistic.


GetDistanceMat.m
to get the distance matrix. It takes a long time with elastic metrics and also large memory of the computer.
You may want to get it step by step.
